Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-10 cordillera-cycle.tex
Restructure atmosphere forcing for readability.

Atmospheric forcing of the model consists of a present-day monthly climatology
$\{T_{m0}, P_{m0}\}$, modified by temperature-offset time series
${\Delta}T_{TS}$ and a lapse-rate correction ${\Delta}T_{LR}$.
\begin{align}
    T_m(t, x, y) &= T_{m0}(x, y) + {\Delta}T_{TS}(t) + {\Delta}T_{LR}(t, x, y) \\
    P_m(t, x, y) &= P_{m0}(x, y)
\end{align}

The present-day climatology $\{T_{m0}, P_{m0}\}$ was computed from the
near-surface air temperature and precipitation rate fields of the NARR over
the period 1979--2000 (Fig.~\ref{fig:atm}). Modern climate of the North
American Cordillera is characterised by strong geographic variations of
seasonality, timing of the precipitation peak and daily temperature variability.
Our choice of data from the NARR is motivated by the need for an accurate,
high-resolution precipitation forcing, as outlined in a previous sensitivity
study \aref.

Temperature offset time-series ${\Delta}T_{TS}$ are derived from proxy records from
the Greenland Ice Core Project (GRIP), the North Greenland Ice Core Project
(NGRIP), the European Project for Ice Coring in Antarctica (EPICA), the Lake
Vostok ice core, and Ocean Drilling Program (ODP) sites 1012 and 1020, both
located off the Californian shore \aref. Palaeo-temperatures from the GRIP and
NGRIP records were calculated using a quadratic equation \aref:
\begin{equation}
    {\Delta}T_{TS}(t) = -11.88[\chem{\delta^{18}O(t)}-\chem{\delta^{18}O}(0)]
                        -0.1925[\chem{\delta^{18}O(t)}^2-\chem{\delta^{18}O}(0)^2]
\end{equation}
while the temperature reconstructions from the Antarctic and oceanic cores were
provided by their authors \aref. All records were scaled linearly in
order to simulate realistic and comparable ice extents at the last
glacial maximum (Table~\ref{tab:records}, Fig.~\ref{fig:timeseries}).

Prior to surface mass balance computation, the model dynamically applies
a~lapse-rate correction to surface air temperature,
\begin{align}
    {\Delta}T_{LR}(t, x, y) &= -\gamma [z_{s}(t, x, y)-z_{ref}] \\
                            &= -\gamma [h(t, x, y)+z_{b}(t, x, y)-z_{ref}]
\end{align}

This correction accounts for the evolution of ice thickness $h$ on the one
hand, and for differences between the the ice flow model basal topography
$z_{b}$ and the climate forcing reference topography $z_{ref}$ on the other
hand. All simulations use an annual temperature lapse rate of
$\gamma = 6\,\unit{{\degree}C\,km^{-1}}$.

For all records, we firstly run a low-resolution simulation, using
a~10\,\unit{km}, 300 by 150~points horizontal grid for SSA and SIA velocities,
up to 51~enthalpy layers in the ice, and 31~temperature layers in the bedrock.
For the GRIP and EPICA records, we then run a second simulation using a higher
horizontal resolution of 6\,\unit{km}, up to 101~ice layers, and 61~bedrock
layers.

>>>>

Atmospheric forcing of the model consists of a present-day monthly climatology
$\{T_{m0}, P_{m0}\}$, modified by a lapse-rate correction ${\Delta}T_{LR}$ and
temperature-offset time series ${\Delta}T_{TS}$.
\begin{align}
    T_m(t, x, y) &= T_{m0}(x, y) + {\Delta}T_{LR}(t) + {\Delta}T_{TS}(t, x, y) \\
    P_m(t, x, y) &= P_{m0}(x, y)
\end{align}

The present-day climatology $\{T_{m0}, P_{m0}\}$ was computed from
near-surface air temperature and precipitation rate fields of the NARR over
the period 1979--2000 (Fig.~\ref{fig:atm}). Modern climate of the North
American Cordillera is characterised by strong geographic variations of
seasonality, timing of the precipitation peak and daily temperature variability.
Our choice of data from the NARR is motivated by the need for an accurate,
high-resolution precipitation forcing, as outlined in a previous sensitivity
study \aref.

The lapse-rate correction ${\Delta}T_{LR}$ is computed using the NARR surface
geopotential height a reference topography $z_{ref}$,
\begin{align}
    {\Delta}T_{LR}(t, x, y) &= -\gamma [z_{s}(t, x, y)-z_{ref}] \\
                            &= -\gamma [h(t, x, y)+z_{b}(t, x, y)-z_{ref}],
\end{align}

thus accounting for the evolution of ice thickness $h$ on the one hand, and
for differences between the the ice flow model basal topography $z_{b}$ and the
NARR topography $z_{ref}$ on the other hand. All simulations use an annual temperature lapse rate of $\gamma = 6\,\unit{{\degree}C\,km^{-1}}$.

Temperature offset time-series ${\Delta}T_{TS}$ are derived from proxy records from
the Greenland Ice Core Project (GRIP), the North Greenland Ice Core Project
(NGRIP), the European Project for Ice Coring in Antarctica (EPICA), the Lake
Vostok ice core, and Ocean Drilling Program (ODP) sites 1012 and 1020, both
located off the Californian shore \aref. Palaeo-temperatures from the GRIP and
NGRIP records were calculated using a quadratic equation \aref,
\begin{equation}
    {\Delta}T_{TS}(t) = -11.88[\chem{\delta^{18}O(t)}-\chem{\delta^{18}O}(0)]
                        -0.1925[\chem{\delta^{18}O(t)}^2-\chem{\delta^{18}O}(0)^2],
\end{equation}
while the temperature reconstructions from the Antarctic and oceanic cores were
provided by their authors \aref. All records were scaled linearly in
order to simulate realistic and comparable ice extents at the last
glacial maximum (Table~\ref{tab:records}, Fig.~\ref{fig:timeseries}).

For all proxy records, we firstly run a low-resolution simulation, using
a~10\,\unit{km}, 300 by 150~points horizontal grid for SSA and SIA velocities,
up to 51~enthalpy layers in the ice, and 31~temperature layers in the bedrock.
For the GRIP and EPICA records, we then run a second simulation using a higher
horizontal resolution of 6\,\unit{km}, up to 101~ice layers, and 61~bedrock
layers.
